---
title: "13_12_2017"
author: "John D."
date: "December 13, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(modelr)
library(broom)
library(ISLR)
library(MASS)
library(dplyr)
library(car)
library(ISLR)
library(ggplot2)
```

## 3.6.3 Multiple Linear Regression

```{r}

lm.fit <- lm(medv~lstat+age ,data=Boston )
summary (lm.fit)

lm.fit <- lm(medv~.,data=Boston)
summary (lm.fit)
summary(lm.fit)$r.sq
summary(lm.fit)$sigma

#install.packages('car')
vif(lm.fit)

lm.fit1 <- lm(medv∼.-age ,data=Boston )
summary (lm.fit1)


lm.fit1 <- update(lm.fit , ∼.-age)
summary (lm.fit1)
```

# Problems

### 1. Describe the null hypotheses to which the p-values given in Table 3.4 correspond. Explain what conclusions you can draw based on these p-values. Your explanation should be phrased in terms of sales, TV, radio, and newspaper, rather than in terms of the coefficients of the linear model.
  
```{r}
advertising <- read.csv("Advertising.csv", row.names = 1)
head(advertising)
lm.fit <- lm(sales ~ ., data = advertising)
summary(lm.fit)
summary(lm.fit)$coefficients
```

The p-values are for the null hypotheses that the predictor has no effect on the response. For TV and radio it is highly unlikely that these predictors do not affect sales. Newspaper on the other hand does not have a significant effect.

### 9. This question involves the use of multiple linear regression on the Auto data set.

  (a) Produce a scatterplot matrix which includes all of the variables in the data set.

```{r}
dat <- Auto
pairs(dat)
```
  
  (b) Compute the matrix of correlations between the variables using the function cor(). You will need to exclude the name variable, which is qualitative.

```{r}
dat2 <- dat[,-9]
head(dat2)
cor(dat2)
symnum(cor(dat2))
```  

  (c) Use the lm() function to perform a multiple linear regression with mpg as the response and all other variables except name as the predictors. Use the summary() function to print the results. Comment on the output. For instance:
    i. Is there a relationship between the predictors and the response?
    ii. Which predictors appear to have a statistically significant relationship to the response?
    iii. What does the coefficient for the year variable suggest?

```{r}
lm.fit <- lm(mpg ~ .-name, data = dat)

#or
#lm.fit <- lm(mpg ~ ., data = dat2)

summary(lm.fit)

```

i. When all predictors are considered together cylinders, horsepower, and acceleration are not found to be significant while the other predictors are.

ii. Displacement, weight, year, and origin all appear to be significant when regressing mpg on all other variables.

iii. The coefficient for year suggests that for each increase 1 year increase in year the mpg of a car increases by .75

### 10. This question should be answered using the Carseats data set.

  (a) Fit a multiple regression model to predict Sales using Price, Urban, and US.
  
```{r}
dat <- Carseats
summary(dat)
lm.fit <- lm(Sales ~ Price+Urban+US, data = dat)
summary(lm.fit)
```

  (b) Provide an interpretation of each coefficient in the model. Be Careful—some of the variables in the model are qualitative!
  
```{r}
summary(lm.fit)$coefficients
```

As price increases the number of sales decreases. If a store is in an urban area sales goes down although this predictor appears to be insignificant based off its p-value. If the store is in the US sales goes up.

  (c) Write out the model in equation form, being careful to handle the qualitative variables properly
  
```{r}
as.formula(
  paste0("Sales ~ ", round(coefficients(lm.fit)[1],2), "", 
    paste(sprintf(" %+.2f*%s ", 
                  coefficients(lm.fit)[-1],  
                  names(coefficients(lm.fit)[-1])),
          collapse="")," + error"
  )
)
```

  (d) For which of the predictors can you reject the null hypothesis H0 : βj = 0?
  
For the predictors Price and US we can reject the null hypothesis

  (e) On the basis of your response to the previous question, fit a smaller model that only uses the predictors for which there is evidence of association with the outcome.
  
```{r}
lm.fit2 <- lm(Sales ~ Price+US, data = dat)
```

  (f) How well do the models in (a) and (e) fit the data?

```{r}
summary(lm.fit)
summary(lm.fit2)
summary(lm.fit)$r.squared-summary(lm.fit2)$r.squared

```

There is not much of a difference between the two models.

  (g) Using the model from (e), obtain 95 % confidence intervals for the coefficient(s).

```{r}
confint(lm.fit2)
```

### 15. This problem involves the Boston data set, which we saw in the lab for this chapter. We will now try to predict per capita crime rate using the other variables in this data set. In other words, per capita crime rate is the response, and the other variables are the predictors.

  (a) For each predictor, fit a simple linear regression model to predict the response. Describe your results. In which of the models is there a statistically significant association between the predictor and the response? Create some plots to back up your assertions.
  
  
  
```{r}
boston.nest <- Boston %>% 
  tidyr::gather(key="predictor",value="value",-crim) %>%
  dplyr::group_by(predictor) %>%
  tidyr::nest()
boston.nest # a set of data frames, one for each predictor
boston.nest$data[[1]]

fitModel <- function(df) lm(crim ~ value, data=df)
boston.nest <- boston.nest%>%
  mutate(model=purrr::map(data,fitModel),
         model.summary=purrr::map(model,summary))
boston.nest

boston_glance <- boston.nest %>%
  mutate(glance=purrr::map(model,glance)) %>%
  tidyr::unnest(glance,.drop = TRUE)
boston_glance
```

  (b) Fit a multiple regression model to predict the response using all of the predictors. Describe your results. For which predictors can we reject the null hypothesis H0 : βj = 0?
  
```{r}
# Cut off of .01
lm.fit <- lm(crim ~ ., data = Boston)
coeffs <- summary(lm.fit)$coefficients[-1,]
rejectable <- coeffs[which(coeffs[,4] < .01),]
rejectable
```

The null hypothesis for dis, rad, and medv can be rejected.

  (c) How do your results from (a) compare to your results from (b)? Create a plot displaying the univariate regression coefficients from (a) on the x-axis, and the multiple regression coefficients from (b) on the y-axis. That is, each predictor is displayed as a single point in the plot. Its coefficient in a simple linear regression model is shown on the x-axis, and its coefficient estimate in the multiple linear regression model is shown on the y-axis.
  
```{r}
all <- boston.nest[,c(1,3)]
all$model_1_coef <- sapply(1:nrow(all), function(x) all$model[[x]][[1]][[2]])
all$model <- NULL
all$model_2_coef <- summary(lm.fit)$coefficients[-1,1]
all

ggplot(data = all, aes(x = model_1_coef, y = model_2_coef)) +
  geom_point() +
  geom_text(aes(label=predictor),hjust=0, vjust=-.5) +
  geom_abline()

ggplot(data = subset(all, subset = predictor != "nox"), aes(x = model_1_coef, y = model_2_coef)) +
  geom_point() +
  geom_text(aes(label=predictor),hjust=0, vjust=-.5) +
  geom_abline()
```

  (d) Is there evidence of non-linear association between any of the predictors and the response? To answer this question, for each predictor X, fit a model of the form: Y = β0 + β1X + β2X2 + β3X3 + e
  
```{r}
boston2 <- boston.nest[,1:2]

polyFitModel <- function(df) lm(crim ~ value + value**2 + value**3, data=df)
boston2 <- boston2%>%
  mutate(model=purrr::map(data,polyFitModel),
         model.summary=purrr::map(model,summary))
boston2

boston2_glance <- boston2 %>%
  mutate(glance=purrr::map(model,glance)) %>%
  tidyr::unnest(glance,.drop = TRUE)
boston2_glance
```