---
title: "2018_05_18"
author: "John D."
date: "May 18, 2018"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 10 Unsupervised Learning

## 10.4 Lab 1: Principal Components Analysis

```{r}
states <- row.names(USArrests)
states

names(USArrests)
apply(USArrests, 2, mean)
apply(USArrests, 2, var)

pr.out <- prcomp(USArrests, scale = T)
names(pr.out)

pr.out$center
pr.out$scale
pr.out$rotation

dim(pr.out$x)
head(pr.out$x)

biplot(pr.out, scale = 0, cex = .5)

pr.out$rotation <- -pr.out$rotation
pr.out$x <- -pr.out$x
biplot(pr.out, scale = 0, cex = .5)

pr.out$sdev
pr.var <- pr.out$sdev^2
pr.var

pve <- pr.var/sum(pr.var)
pve

plot(pve, xlab = "Principal Component", ylab = "Proportion of Variance Explained", ylim = c(0,1), type = "b")
plot(cumsum(pve), xlab = "Principal Component", ylab = "Cumulative Proportion of Variance Explained", ylim = c(0,1), type = "b")

a <- c(1,2,8,-3)
cumsum(a)
```

## 10.7 Exercises

### Conceptual

#### 6. A researcher collects expression measurements for 1,000 genes in 100 tissue samples. The data can be written as a 1, 000 × 100 matrix, which we call X, in which each row represents a gene and each column a tissue sample. Each tissue sample was processed on a different day, and the columns of X are ordered so that the samples that were processed earliest are on the left, and the samples that were processed later are on the right. The tissue samples belong to two groups: control (C) and treatment (T). The C and T samples were processed in a random order across the days. The researcher wishes to determine whether each gene’s expression measurements differ between the treatment and control groups.
#### As a pre-analysis (before comparing T versus C), the researcher performs a principal component analysis of the data, and finds that the first principal component (a vector of length 100) has a strong linear trend from left to right, and explains 10 % of the variation. The researcher now remembers that each patient sample was run on one of two machines, A and B, and machine A was used more often in the earlier times while B was used more often later. The researcher has a record of which sample was run on which machine

  (a) Explain what it means that the first principal component “explains
10% of the variation”  
  The first principal component contains 10% of the overall variance. 90% of the variance is not contained by the first principal component.
  
  (b) The researcher decides to replace the (j, i)th element of X with xji − φj1zi1 where zi1 is the ith score, and φj1 is the jth loading, for the first principal component. He will then perform a two-sample t-test on each gene in this new data set in order to determine whether its expression differs between the two conditions. Critique this idea, and suggest a better approach. (The principal component analysis is performed on XT ). 
  Should probably factor in which machine was used, A or B.
  
  (c) Design and run a small simulation experiment to demonstrate the superiority of your idea.  
  
```{r}
set.seed(123)
dat <- matrix(rnorm(100 * 1000), ncol = 100)
dat[1, ] <- seq(from = -20, to =20, length.out = 100) # adding replacement
pr.out <- prcomp(scale(dat))

summary(pr.out)
summary(pr.out)$importance[,1]

dat <- rbind(dat, c(rep(-10, 50), rep(10, 50)))
pr.out <- prcomp(scale(dat))
summary(pr.out)$importance[, 1]
```
  Went up a bit.
  
### Applied

#### 8. In Section 10.2.3, a formula for calculating PVE was given in Equation 10.8. We also saw that the PVE can be obtained using the sdev output of the prcomp() function. On the USArrests data, calculate PVE in two ways:

  (a) Using the sdev output of the prcomp() function, as was done in
Section 10.2.3  
```{r}
pr.out <- prcomp(USArrests, scale = T)
pr.out$sdev
pr.var <- pr.out$sdev^2
pve <- pr.var/sum(pr.var)
pve
```
  (b) By applying Equation 10.8 directly. That is, use the prcomp() function to compute the principal component loadings. Then, use those loadings in Equation 10.8 to obtain the PVE.  
```{r}
pr.out <- prcomp(USArrests, scale = T)
scaled.USArrests <- scale(USArrests)
top.sums <- apply((scaled.USArrests %*% pr.out$rotation)^2, 2, sum) # Top half of equation, each pc variance
bottom.sum <- sum(apply(scaled.USArrests^2, 2, sum)) # Bottom half of equation, total variance
top.sums/bottom.sum
```

#### 10. In this problem, you will generate simulated data, and then perform PCA and K-means clustering on the data.

  (a) Generate a simulated data set with 20 observations in each of three classes (i.e. 60 observations total), and 50 variables.  
```{r}
set.seed(321)
dat <- matrix(rnorm(20 * 3 * 50, mean = 0), ncol = 50)
dat[1:20,1] <- dat[1:20,1] + 3
dat[21:40,2] <- dat[21:40,2] - 3
dat[41:60,3] <- dat[41:60,3] + 4
groupings <- c(rep(1, 20), rep(2, 20), rep(3, 20))
```

  (b) Perform PCA on the 60 observations and plot the first two principal component score vectors. Use a different color to indicate the observations in each of the three classes. If the three classes appear separated in this plot, then continue on to part (c). If not, then return to part (a) and modify the simulation so that there is greater separation between the three classes. Do not continue to part (c) until the three classes show at least some separation in the first two principal component score vectors.
  
```{r}
pr.out <- prcomp(dat)
plot(pr.out$x[, 1:2], col = groupings, xlab = "Principal Component 1 Score", ylab = "Principal Component 2 Score", pch = 19)
```

