---
title: "2017_12_06"
author: "John D."
date: "December 6, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 3.6 Lab: Linear Regression
```{r}
library(MASS)
library(ISLR)
```

## 3.6.2 Simple Linear Regression

```{r}
#fix(Boston)
names(Boston)
lm.fit <- lm(medv∼lstat , data=Boston)
attach(Boston)
lm.fit
summary(lm.fit)
names(lm.fit)
confint(lm.fit)

predict(lm.fit ,data.frame(lstat=c(5,10 ,15)),
interval ="confidence")

predict(lm.fit ,data.frame(lstat=c(5,10 ,15)),
interval ="prediction")

plot(lstat ,medv)
abline(lm.fit)

abline (lm.fit ,lwd =3)
abline (lm.fit ,lwd=3,col ="red")
plot(lstat ,medv ,col="red")
plot(lstat ,medv ,pch =20)
plot(lstat ,medv ,pch ="+")
plot(1:20,1:20,pch =1:20)
par(mfrow=c(2,2))
plot(lm.fit)

plot(predict (lm.fit), residuals (lm.fit))
plot(predict (lm.fit), rstudent (lm.fit))
plot(hatvalues (lm.fit))
which.max(hatvalues (lm.fit))
```

## Problem 8

8. This question involves the use of simple linear regression on the Auto data set.
  (a) Use the lm() function to perform a simple linear regression with mpg as the response and horsepower as the predictor. Use the summary() function to print the results. Comment on the output.
    i. Is there a relationship between the predictor and the response?
    Yes
    
    ii. How strong is the relationship between the predictor and the response?
    Pretty strong
    
    iii. Is the relationship between the predictor and the response positive or negative?
    Negative
    
    iv. What is the predicted mpg associated with a horsepower of 98? What are the associated 95 % confidence and prediction intervals
    
```{r}
fit <- lm(mpg~horsepower, data = Auto)
summary(fit)

predict(fit ,data.frame(horsepower=98),
interval ="prediction")

predict(fit ,data.frame(horsepower=98),
interval ="confidence")
```

(b) Plot the response and the predictor. Use the abline() function to display the least squares regression line.)

```{r}
par(mfrow=c(1,1))
attach(Auto)
plot(mpg~horsepower)
abline(fit)
```

(c) Use the plot() function to produce diagnostic plots of the least squares regression fit. Comment on any problems you see with the fit.

```{r}
par(mfrow=c(2,2))
plot(fit)

```

## Problem 13

In this exercise you will create some simulated data and will fit simple linear regression models to it. Make sure to use set.seed(1) prior to starting part (a) to ensure consistent results.

(a) Using the rnorm() function, create a vector, x, containing 100 observations drawn from a N(0, 1) distribution. This represents a feature, X.

```{r}
set.seed(1)
x <- rnorm(100,0,1)
```

(b) Using the rnorm() function, create a vector, eps, containing 100 observations drawn from a N(0, 0.25) distribution i.e. a normal distribution with mean zero and variance 0.25.

```{r}
eps <- rnorm(100,0,.25)
```

(c) Using x and eps, generate a vector y according to the model Y = −1+0.5X + eps. What is the length of the vector y? What are the values of βo and β1 in this linear model?

```{r}
y = -1+0.5*x+eps
length(y)
#βo = -1 
#β1 = .5
```

(d) Create a scatterplot displaying the relationship between x and y. Comment on what you observe.

```{r}
par(mfrow=c(1,1))
plot(y~x)
```

Linear relationship

(e) Fit a least squares linear model to predict y using x. Comment on the model obtained. How do β^o and β^1 compare to βo and β1?

```{r}
fit <- lm(y~x)
summary(fit)
```

Almost identical

(f)Display the least squares line on the scatterplot obtained in (d).Draw the population regression line on the plot, in a different color. Use the legend() command to create an appropriate legend

```{r}
plot(y~x); abline(fit, col ="red") 
legend("bottomright", c("Regression line"), lwd=1, col="red",bty ="n")
```

(g)Now fit a polynomial regression model that predicts y using x and x^2. Is there evidence that the quadratic term improves the model fit? Explain your answer

```{r}
fit2 = lm(y~poly(x,2))
summary(fit2)
```

Not significant

(h)Repeat (a)–(f) after modifying the data generation process in such a way that there is less noise in the data. The model (part-c) should remain the same. You can do this by decreasing the variance of the normal distribution used to generate the error term in (b). Describe your results.

```{r}
x = rnorm(100, mean= 0, sd =1)
eps = rnorm(100, mean =0, sd = 0.05)
y = -1+0.5*x+eps
fit3 = lm(y~x)
summary(fit3)
```

(j)What are the confidence intervals for βo and β1 based on the original data set, the noisier data set, and the less noisy data set? Comment on your results.
```{r}
confint(fit)
confint(fit2)
confint(fit3)
```

## Problem 15

15. This problem involves the Boston data set, which we saw in the lab for this chapter. We will now try to predict per capita crime rate using the other variables in this data set. In other words, per capita crime rate is the response, and the other variables are the predictors.

(a)For each predictor, fit a simple linear regression model to predict the response. Describe your results. In which of the models is there a statistically significant association between the predictor and the response? Create some plots to back up your assertions.

```{r}

results <- lapply(Boston[,2:ncol(Boston)], function(x) lm(crim~x, data = Boston))
for(i in 1:13){
  print(names(results[i]))
  print(summary(results[[i]]))
  par(mfrow=c(2,2))
  plot(results[[i]])
}
```
