---
title: "2018_05_11"
author: "John D."
date: "May 11, 2018"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load data

```{r}
library(ISLR)
library(ggplot2)
library(dplyr)
library(e1071)
library(ROCR)
library(reshape2)
```

# 9.6 Lab: Support Vector Machines
## 9.6.1 Support Vector Classifier
```{r}
set.seed(1)
x <- matrix(rnorm(20*2), ncol=2)
y <- c(rep(-1,10), rep(1,10))
x[y==1,] <- x[y==1,] + 1
plot(x, col=(3-y))

dat <- data.frame(x=x, y=as.factor(y))
svmfit <- svm(y ~., data=dat, kernel = "linear", cost = 10, scale = FALSE)

plot(svmfit, dat)

svmfit$index
summary(svmfit)

svmfit <- svm(y ~ ., data=dat, kernel = "linear", cost = 0.1, scale=FALSE)
plot(svmfit, dat)
svmfit$index

set.seed(1)
tune.out <- tune(svm, y ~ ., data=dat, kernel = "linear", ranges=list(cost=c(0.001, 0.01, 0.1, 1, 5, 10, 100)))

summary(tune.out)

bestmod <- tune.out$best.model
summary(bestmod)

xtest <- matrix(rnorm(20*2), ncol=2)
ytest <- sample (c(-1,1), 20, rep=TRUE)
xtest[ytest==1,] <-  xtest[ytest==1,] + 1
testdat <- data.frame(x = xtest, y = as.factor(ytest))

ypred <- predict(bestmod, testdat)
table(predict = ypred, truth = testdat$y)

svmfit <- svm(y ~ ., data=dat, kernel = "linear", cost = .01, scale=FALSE)
ypred <- predict(svmfit, testdat)
table(predict = ypred, truth = testdat$y)

x[y==1,] <- x[y==1,] + 0.5
plot(x, col = (y+3)/2, pch = 19)

dat <- data.frame(x = x, y = as.factor(y))
svmfit <- svm(y ~ ., data=dat, kernel = "linear", cost = 1e5)
summary(svmfit)
plot(svmfit, dat)

svmfit <- svm(y ~ ., data=dat, kernel = "linear", cost = 1)
summary(svmfit)
plot(svmfit, dat)
```

## 9.6.2 Support Vector Machine

```{r}
set.seed(1)
x=matrix(rnorm (200*2) , ncol=2)
x[1:100,]=x[1:100,]+2
x[101:150,] = x[101:150,]-2
y=c(rep(1,150) ,rep(2,50))
dat=data.frame(x=x,y=as.factor(y))

plot(x, col=y)

train=sample(200,100)
svmfit=svm(y∼., data=dat[train ,], kernel ="radial", gamma=1,cost=1)
plot(svmfit , dat[train ,])

summary(svmfit)

svmfit=svm(y∼., data=dat[train ,], kernel ="radial",gamma=1,cost=1e5)
plot(svmfit ,dat[train ,])

set.seed(1)
tune.out=tune(svm , y∼., data=dat[train ,], kernel="radial",ranges=list(cost=c(0.1,1,10,100,1000),gamma=c(0.5,1,2,3,4) ))
summary(tune.out)

table(true=dat[-train ,"y"], pred=predict(tune.out$best.model ,newdata =dat[-train ,]))
```

## 9.6.3 ROC Curves

```{r}
rocplot =function (pred, truth, ...){
  predob = prediction(pred, truth)
  perf = performance(predob, "tpr", "fpr")
  plot(perf,...)}

svmfit.opt=svm(y∼., data=dat[train ,], kernel ="radial", gamma=2, cost=1, decision.values =T)
fitted =attributes(predict(svmfit.opt ,dat[train ,], decision.values=TRUE))$decision.values

par(mfrow=c(1,2))
rocplot(fitted ,dat[train ,"y"], main="Training Data")

svmfit.flex=svm(y∼., data=dat[train ,], kernel ="radial",gamma=50, cost=1, decision.values =T)
fitted=attributes(predict (svmfit.flex ,dat[train ,], decision.values=T))$decision.values
rocplot(fitted ,dat[train ,"y"],add=T,col="red ")

fitted=attributes(predict (svmfit.opt ,dat[-train ,], decision.values=T))$decision.values
rocplot (fitted ,dat[-train ,"y"], main="Test Data")
fitted=attributes (predict (svmfit.flex ,dat[- train ,], decision.values=T))$decision.values
rocplot(fitted ,dat[-train ,"y"],add=T,col="red")
```

## 9.6.4 SVM with Multiple Classes

```{r}
set.seed(1)
x=rbind(x, matrix(rnorm (50*2) , ncol=2))
y=c(y, rep(0,50))
x[y==0,2] = x[y==0 ,2]+2
dat=data.frame(x=x, y=as.factor(y))
par(mfrow=c(1,1))
plot(x,col=(y+1))

svmfit=svm(y∼., data=dat , kernel ="radial", cost=10, gamma =1)
plot(svmfit, dat)
```

## 9.6.5 Application to Gene Expression Data

```{r}
names(Khan)
dim(Khan$xtrain)
dim(Khan$xtest)
length(Khan$ytrain)
length(Khan$ytest)

table(Khan$ytrain)
table(Khan$ytest)

dat=data.frame(x=Khan$xtrain , y=as.factor(Khan$ytrain ))
out=svm(y∼., data=dat, kernel ="linear",cost=10)
summary(out)

table(out$fitted , dat$y)

dat.te=data.frame(x=Khan$xtest , y=as.factor(Khan$ytest ))
pred.te=predict(out, newdata=dat.te)
table(pred.te, dat.te$y)
```

## 4. Generate a simulated two-class data set with 100 observations and two features in which there is a visible but non-linear separation between the two classes. Show that in this setting, a support vector machine with a polynomial kernel (with degree greater than 1) or a radial kernel will outperform a support vector classifier on the training data. Which technique performs best on the test data? Make plots and report training and test error rates in order to back up your assertions.

```{r}
# Make data
set.seed(123)
x <- matrix(rnorm (100*2) , ncol=2)
y <- ifelse(x[,1] > -.1 & x[,1] < .75 & x[,2] > 0 & x[,2] < 1.5, -1, 1)
dat=data.frame(x=x, y=as.factor(y))
par(mfrow=c(1,1))
plot(x,col=(y+3))

# Create training set
train <- sample(100, 75)

# Linear
linear.tune<- tune(svm, y ~ ., data=dat[train,], kernel = "linear", ranges=list(cost=c(0.01, 0.1, 1, 5, 10)))
summary(linear.tune)
linear.bestmod <- linear.tune$best.model
summary(linear.bestmod)

table(predict = predict(linear.bestmod, dat[train,]), truth = dat[train,3])
100 - sum(predict(linear.bestmod, dat[train,]) == dat[train,3])/length(dat[train,3])*100
# 18.67% training error

table(predict = predict(linear.bestmod, dat[-train,]), truth = dat[-train,3])
100 - sum(predict(linear.bestmod, dat[-train,]) == dat[-train,3])/length(dat[-train,3])*100
# 12% test error

plot(linear.bestmod, dat[train,])
plot(linear.bestmod, dat[-train,])

# Radial
radial.tune<- tune(svm, y ~ ., data=dat[train,], kernel="radial",ranges=list(cost=c(0.1,1,10,100,1000),gamma=c(0.5,1,2,3,4)))
summary(radial.tune)
radial.bestmod <- radial.tune$best.model
summary(radial.bestmod)

table(predict = predict(radial.bestmod, dat[train,]), truth = dat[train,3])
100 - sum(predict(radial.bestmod, dat[train,]) == dat[train,3])/length(dat[train,3])*100
# 0% training error

table(predict = predict(radial.bestmod, dat[-train,]), truth = dat[-train,3])
100 - sum(predict(radial.bestmod, dat[-train,]) == dat[-train,3])/length(dat[-train,3])*100
# 0% test error

plot(radial.bestmod, dat[train,])
plot(radial.bestmod, dat[-train,])
```

## 5. We have seen that we can fit an SVM with a non-linear kernel in order to perform classification using a non-linear decision boundary. We will now see that we can also obtain a non-linear decision boundary by performing logistic regression using non-linear transformations of the features.

  (a) Generate a data set with n = 500 and p = 2, such that the observations belong to two classes with a quadratic decision boundary between them. For instance, you can do this as follows:

```{r}
x1 <- runif(500)-0.5
x2 <- runif(500)-0.5
y <- 1*(x1^2-x2^2 > 0)
```

  (b) Plot the observations, colored according to their class labels. Your plot should display X1 on the x-axis, and X2 on the yaxis.
  
```{r}
plot(x1,x2, col = c(y + 2))
```


  (c) Fit a logistic regression model to the data, using X1 and X2 as predictors.

```{r}

train <- sample(500,400)
dat <- data.frame(x1,x2, y = factor(y))
glm.fits=glm(y ∼ ., data=dat[train,] ,family=binomial)
summary(glm.fits)
```

  (d) Apply this model to the training data in order to obtain a predicted class label for each training observation. Plot the observations, colored according to the predicted class labels. The decision boundary should be linear.

```{r}
glm.probs <- predict(glm.fits, type="response")
contrasts(dat$y)
glm.pred <- rep(0 ,400)
glm.pred[glm.probs >.5] <- 1
table(glm.pred, dat[train,]$y)
mean(glm.pred == dat[train,]$y)

plot(dat$x1[train], dat$x2[train], col = c(glm.pred + 2))
```


  (e) Now fit a logistic regression model to the data using non-linear functions of X1 and X2 as predictors (e.g. X2 1 , X1×X2, log(X2), and so forth).
```{r}
glm.fits=glm(y ∼ x1 + x2 + I(x1^2) + I(x2^2) + x1*x2, data=dat[train,] ,family=binomial)
summary(glm.fits)
```

  (f) Apply this model to the training data in order to obtain a predicted class label for each training observation. Plot the observations, colored according to the predicted class labels. The decision boundary should be obviously non-linear. If it is not, then repeat (a)-(e) until you come up with an example in which the predicted class labels are obviously non-linear.

```{r}
glm.probs <- predict(glm.fits, type="response")
glm.pred <- rep(0 ,400)
glm.pred[glm.probs >.5] <- 1
table(glm.pred, dat[train,]$y)
mean(glm.pred == dat[train,]$y)

plot(dat$x1[train], dat$x2[train], col = c(glm.pred + 2))
```

  (g) Fit a support vector classifier to the data with X1 and X2 as predictors. Obtain a class prediction for each training observation. Plot the observations, colored according to the predicted class labels.

```{r}
svc.tune  <- tune(svm, y ∼ . , data=dat[train,], kernel="linear", ranges=list(cost=c(0.001, 0.01, 0.1, 1, 5, 10, 100)))
svc.fit <- svc.tune$best.model

table(predict = predict(svc.fit, dat[train,]), truth = dat[train,3])
100 - sum(predict(svc.fit, dat[train,]) == dat[train,3])/length(dat[train,3])*100

plot(dat$x1, dat$x2, col = c(predict(svc.fit, dat[train,])))
```

  (h) Fit a SVM using a non-linear kernel to the data. Obtain a class prediction for each training observation. Plot the observations, colored according to the predicted class labels.

```{r}
# Radial
svm.tune  <- tune(svm, y ∼ . , data=dat[train,], kernel="radial",ranges=list(cost=c(0.1,1,10,100,1000),gamma=c(0.5,1,2,3,4)))
svm.fit <- svm.tune$best.model

table(predict = predict(svm.fit, dat[train,]), truth = dat[train,3])
100 - sum(predict(svm.fit, dat[train,]) == dat[train,3])/length(dat[train,3])*100

plot(dat$x1[train], dat$x2[train], col = c(predict(svm.fit, dat[train,])))

# Polynomial
svm.tune  <- tune(svm, y ∼ . , data=dat[train,], kernel="polynomial",ranges=list(cost=c(0.1,1,10,100,1000),degree=c(1,2,3)))
svm.fit <- svm.tune$best.model

table(predict = predict(svm.fit, dat[train,]), truth = dat[train,3])
100 - sum(predict(svm.fit, dat[train,]) == dat[train,3])/length(dat[train,3])*100

plot(dat$x1[train], dat$x2[train], col = c(predict(svm.fit, dat[train,])))
```


  (i) Comment on your results.

Similar

# #7. In this problem, you will use support vector approaches in order to predict whether a given car gets high or low gas mileage based on the Auto data set.

(a) Create a binary variable that takes on a 1 for cars with gas mileage above the median, and a 0 for cars with gas mileage below the median.

```{r}
summary(Auto)
Auto$efficiency <- as.factor(ifelse(Auto$mpg > median(Auto$mpg), 1, 0))
```

(b) Fit a support vector classifier to the data with various values of cost, in order to predict whether a car gets high or low gas mileage. Report the cross-validation errors associated with different values of this parameter. Comment on your results.

```{r}
set.seed(1)
tune.out <- tune(svm, efficiency ~ ., data=Auto, kernel = "linear", ranges=list(cost=c(0.001, 0.01, 0.1, 1, 5, 10, 100)))

summary(tune.out)
summary(tune.out)$best.performance

linear.bestmod <- tune.out$best.model
summary(linear.bestmod)
```

Best cost was somewhere around 1 and 5.

  (c) Now repeat (b), this time using SVMs with radial and polynomial basis kernels, with different values of gamma and degree and cost. Comment on your results.

```{r}
#Radial
tune.out <- tune(svm, efficiency ~ ., data=Auto, kernel = "radial", ranges=list(cost=c(0.1,1,10,100,1000),gamma=c(0.5,1,2,3,4)))

summary(tune.out)
summary(tune.out)$best.performance

radial.bestmod <- tune.out$best.model
summary(radial.bestmod)

tune.out <- tune(svm, efficiency ~ ., data=Auto, kernel = "polynomial", ranges=list(cost=c(0.1,1,10,100,1000),degree=c(1,2,3)))

summary(tune.out)
summary(tune.out)$best.performance

poly.bestmod <- tune.out$best.model
summary(poly.bestmod)

```

  (d) Make some plots to back up your assertions in (b) and (c).
  
```{r}
dat <- data.frame(Observation = 1:nrow(Auto), Truth = Auto$efficiency, linear = predict(linear.bestmod, Auto),
                  radial = predict(radial.bestmod, Auto), polynomial = predict(poly.bestmod, Auto))
par(mfrow=c(2,2))
plot(dat$Observation,dat$Truth, col = 1)
plot(dat$Observation,dat$linear, col = 2)
plot(dat$Observation,dat$radial, col = 3)
plot(dat$Observation,dat$polynomial, col = 4)
```

## 8. This problem involves the OJ data set which is part of the ISLR package.

(a) Create a training set containing a random sample of 800 observations, and a test set containing the remaining observations.

```{r}
summary(OJ)
set.seed(3251)
train <- sample(1:nrow(OJ), 800, replace = F)
dat.train <- OJ[train,]
dat.test <- OJ[-train,]
```

(b) Fit a support vector classifier to the training data using cost=0.01, with Purchase as the response and the other variables as predictors. Use the summary() function to produce summary statistics, and describe the results obtained.

```{r}
svmfit <- svm(Purchase ~ ., data = dat.train, kernel = "linear", cost = 0.01, scale = TRUE)
summary(svmfit)
```
About 218 support vectors from each class

(c) What are the training and test error rates?

```{r}
pred.train <- predict(svmfit, dat.train)
table(predict = pred.train, truth = dat.train$Purchase)
100 - sum(pred.train == dat.train$Purchase)/length(dat.train$Purchase)*100
# 16.5% training error

pred.test <- predict(svmfit, dat.test)
table(predict = pred.test, truth = dat.test$Purchase)
100 - sum(pred.test == dat.test$Purchase)/length(dat.test$Purchase)*100
# 16.7% test error
```


(d) Use the tune() function to select an optimal cost. Consider values in the range 0.01 to 10.

```{r}
set.seed(567)
tune.out <- tune(svm, Purchase ~ ., data=dat.train, kernel = "linear", ranges=list(cost=c(0.01, 0.1, 1, 5, 10)), scale = T)

summary(tune.out)

bestmod <- tune.out$best.model
summary(bestmod)
```

(e) Compute the training and test error rates using this new value for cost.

```{r}
pred.train <- predict(bestmod, dat.train)
table(predict = pred.train, truth = dat.train$Purchase)
100 - sum(pred.train == dat.train$Purchase)/length(dat.train$Purchase)*100
# 16.5% training error

pred.test <- predict(bestmod, dat.test)
table(predict = pred.test, truth = dat.test$Purchase)
100 - sum(pred.test == dat.test$Purchase)/length(dat.test$Purchase)*100
# 16.7% test error
```

The predicted best cost is the same as the previous steps

(f) Repeat parts (b) through (e) using a support vector machine with a radial kernel. Use the default value for gamma.

```{r}
svmfit <- svm(Purchase ~ ., data = dat.train, kernel = "radial", cost = .01, scale = T)
summary(svmfit)

pred.train <- predict(svmfit, dat.train)
table(predict = pred.train, truth = dat.train$Purchase)
100 - sum(pred.train == dat.train$Purchase)/length(dat.train$Purchase)*100
# 37.88% training error

pred.test <- predict(svmfit, dat.test)
table(predict = pred.test, truth = dat.test$Purchase)
100 - sum(pred.test == dat.test$Purchase)/length(dat.test$Purchase)*100
# 42.22% test error

tune.out <- tune(svm, Purchase ~ ., data=dat.train, kernel = "radial", ranges=list(cost=c(0.01, 0.1, 1, 5, 10)), scale = T)

summary(tune.out)

bestmod <- tune.out$best.model
summary(bestmod)

pred.train <- predict(bestmod, dat.train)
table(predict = pred.train, truth = dat.train$Purchase)
100 - sum(pred.train == dat.train$Purchase)/length(dat.train$Purchase)*100
# 13.63% training error

pred.test <- predict(bestmod, dat.test)
table(predict = pred.test, truth = dat.test$Purchase)
100 - sum(pred.test == dat.test$Purchase)/length(dat.test$Purchase)*100

# 19.63% test error
```

(g) Repeat parts (b) through (e) using a support vector machine with a polynomial kernel. Set degree=2.

```{r}
svmfit <- svm(Purchase ~ ., data = dat.train, kernel = "polynomial", cost = .01, degree = 2, scale = T)
summary(svmfit)

pred.train <- predict(svmfit, dat.train)
table(predict = pred.train, truth = dat.train$Purchase)
100 - sum(pred.train == dat.train$Purchase)/length(dat.train$Purchase)*100
# 37.88% training error

pred.test <- predict(svmfit, dat.test)
table(predict = pred.test, truth = dat.test$Purchase)
100 - sum(pred.test == dat.test$Purchase)/length(dat.test$Purchase)*100
# 42.22% test error

tune.out <- tune(svm, Purchase ~ ., data=dat.train, kernel = "polynomial", ranges=list(cost=c(0.01, 0.1, 1, 5, 10)), degree = 2, scale = T)

summary(tune.out)

bestmod <- tune.out$best.model
summary(bestmod)

pred.train <- predict(bestmod, dat.train)
table(predict = pred.train, truth = dat.train$Purchase)
100 - sum(pred.train == dat.train$Purchase)/length(dat.train$Purchase)*100
# 14.63% training error

pred.test <- predict(bestmod, dat.test)
table(predict = pred.test, truth = dat.test$Purchase)
100 - sum(pred.test == dat.test$Purchase)/length(dat.test$Purchase)*100

# 18.52 test error
```

(h) Overall, which approach seems to give the best results on this data?

The linear model gave the best result